---
title: 'Class 8: Machine Learning'
author: "Melissa Ito"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's start with an example of running the **kmeans()** function

```{r}
 # Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)

# Trying to fix an error
cl <- kmeans (x, 2, nstart = 20)
```

 Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?
  A: 30
  
```{r}
cl$size
```
  
  
Q. What ‘component’ of your result object details
      - cluster size?
      - cluster assignment/membership?
      - cluster center?
      
```{r}
# cluster assignment
cl$cluster
```
```{r}
# centers
cl$centers
```
      
      
Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points

```{r}
kmeans(x, centers = 2, nstart = 20)
```

```{r}
cl <- kmeans (x, 2, nstart = 20)
plot(x, col=cl$cluster)
points(cl$centers, pch = 18, col = "blue", cex = 3)
```

## Hierarchical Clustering

We must give the **hclust()** function a distance matrix not the raw data as input

```{r}
# Distance matrix
d <- dist(x)

# Clustering
hc <- hclust (d)
plot (hc)
abline( h = 12, col = "red")
```

```{r}
cutree(hc, h = 12)
```

```{r}
 # Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

 Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
 
Q. How does this compare to your known 'col' groups?

```{r}
d.2 <- dist (x)
hc.2 <- hclust (d.2)
plot (hc.2)
abline(h = 2.5, col = "red")
abline (h = 2, col = "blue")
```

```{r}
grp2 <-  cutree (hc.2, h = 2.5)
grp3 <-  cutree (hc.2, h = 2)
```

```{r}
grp2
```

```{r}
grp3
```

```{r}
plot(x, col = grp2)
```

```{r}
plot(x, col = grp3)
```

```{r}
table(grp2)
```
```{r}
table(grp3)
```

## Principal Component Analysis (PCA)

We will use the **prcomp()** function for PCA
```{r}
## You can also download this file from the class website!
     mydata <- read.csv("https://tinyurl.com/expression-CSV",
                        row.names=1)
     head(mydata, 10)
```

100 genes in this dataaset
```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

NOTE: **prcomp()** expects genes to be columns and samples to be rows...need to transpose matrix with **t()** function
```{r}
head( t(mydata))
```

Run our PCA analysis on the transpose of our data.
```{r}
pca <- prcomp( t(mydata), scale = TRUE)
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2")
```

Calculate the percent variance captures in each PC
```{r}
## Percent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

head(pca.var.per)
```

```{r}
barplot(pca.var.per, main="Scree Plot",
        xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

## Hands-on Section

```{r}
x <- read.csv("https://bioboot.github.io/bggn213_f17/class-material/UK_foods.csv", row.names = 1)
x
```

Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?
A1. 17 rows and 5 columns
```{r}
nrow(x)
ncol(x)
```

Preview the first 6 rows of data
```{r}
#View(x)
head(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?
A3: change beside = from TRUE to FALSE
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?
A6. N. Ireland has a lot of point concentrated on the lower left corner whereas the other countries are more spread out. If everything was same, the points would all lie directly on diagonal.
```{r}
pairs(x, col=rainbow(10), pch=16)
```

Using PCA to better analyze our data
```{r}
pca <- prcomp( t(x))
summary (pca)
```

Generate plot of PC1 vs PC2
```{r}
# plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
# text(pca$x[,1], pca$x[,2], colnames(x), col = mycols)
```

```{r}
mycols <-  c("orange", "red", "blue", "dark green")
```

## Digging Deeper (variable loadings)

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```
```{r}
## Lets focus on PC2 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```






