---
title: "Class 12: Bioinformatics in drug discovery and design"
author: "Melissa Ito"
date: "5/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for Docking

We will first prepare our HIV-Pr system for drug docking by making a protein only PDB format file (i.e. we will remove water, existing ligand etc.)

```{r}
# install.packages("bio3d")
```

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

```{r}
hiv <- read.pdb(file.name)
hiv
```

```{r}
prot <- atom.select(hiv, "protein", value = TRUE)
write.pdb(prot, file = "1hsg_protein.pdb")
prot
```

```{r}
lig <- atom.select(hiv, "ligand", value = TRUE)
write.pdb(lig, file = "1hsg_ligand.pdb")
lig
```

We will load these into ADT to add hydrogens and make PDBQT files for each.

Now we run autodoc vina with the cmd.

## Process our docking results

```{r}
res <-  read.pdb("all.pdbqt", multi = TRUE)
write.pdb(res, "results.pdb")
```


Calculate root mean square distance (RMSD) between each of the docking results and the known crystal strucure
```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

## Normal Mode analysis for flexible prediction

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```



























