---
title: 'Class 6: R Functions'
author: "Melissa Ito"
date: "4/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About
This is my **class 6** R markdown document with my *code* and notes for the day. 
```{r}
plot (1:10, type = "l", col = "blue")
```


## Reading in files practice

There are 3 files to read. 

```{r}
read.table ("test1.txt", header = TRUE, sep = ",")
```
using defaults from read.csv saves time
```{r}
read.csv ("test1.txt")
```

specify $ symbol separating variables
```{r}
read.table ("test2.txt", header = TRUE, sep = "$")
```

test3.txt is separated by white space not tabs or comma or semicolon therefore use read.table
```{r}
read.table ("test3.txt", header = FALSE)
```

can also use URL instead of downloading into class6 file
```{r}
read.csv ("https://bioboot.github.io/bimm143_S19/class-material/test2.txt", header = TRUE, sep = "$")
```

## Functions

Our first silly function example:

```{r}
add <-  function(x, y = 1) {
  #the body!!
  x + y
}
```

```{r}
add(10)
```

```{r}
add (10, 10)
```

default y = 1 and adds 1 to each x value
specify y = 4 will add 4 to each x value
```{r}
add ( c (1, 2, 2))
add( c (1, 2, 3), 4)
c (1, 2, 3) + 4
```
these functions will result in errors
```{r}
#add (1, 2, 2)
#add (x =1, y = "b")
```
 
## Another example

```{r}
rescale <- function(x) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}
```
```{r}
rescale (1:10)
```
```{r}
# How would you get your function to work here...
rescale( c(1,2,NA,3,10) )
```
```{r}
x <-  c(1,2,NA,3,10) 
rng <-range(x, na.rm = TRUE)
rng
    (x - rng[1]) / (rng[2] - rng[1])
```

```{r}
range (c(1,2,3,10))
```
```{r}
rescale2 <- function(x, na.rm = TRUE) {
   rng <-range(x, na.rm = na.rm)
   (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2( c(1, 3, NA, 10))
```

```{r}
# What should your function do here?
 #rescale( c(1,10,"string") ) 

```
```{r}
# x <- c (1, 10, "string")

if(!is.numeric (x)) {
  stop ("You need to give only numbers please")
}

# range (x, na.rm = TRUE)
# (x - rng[1]) / (rng[2] - rng[1])
```
```{r}
#rescale2 (c (1, 10, "string"))
```


```{r}
 rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
   
    rng <-range(x, na.rm=na.rm)
    print("Hello")
    
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   
   
   print("is it me you are looking for?") 

   if(plot) {
      plot(answer, typ="b", lwd=4)
     print("Please don't sing again...")
   }
   print("I can see it in ...")
   return(answer)
 }
```
```{r}
rescale3 ( c(1, 3, NA, 10))
```
```{r}
rescale3 ( c(1, 3, NA, 10), plot = TRUE)
```
```{r}
rescale3a <- function(x, na.rm=TRUE, plot=FALSE) {
   
    rng <-range(x, na.rm=na.rm)
    print("Hello")
    
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   
    return (answer)
   
   print("is it me you are looking for?") 

   if(plot) {
      plot(answer, typ="b", lwd=4)
     print("Please don't sing again...")
   }
   print("I can see it in ...")
   return(answer)
 }
```
```{r}
# will stop processing once it hits the first return(answer) function
rescale3a (c(1, 2, NA, 10))
```



















