---
title: "Class 7: R functions and packages"
author: "Melissa Ito"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions revisited

We will source a file from online with our functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function

```{r}
rescale( 1:10 )
```

Try the rescale2 function that catches string inputs

```{r}
# rescale2(c(1:10, "string"))
```

# Find missing NA values in two vectors

Start with a simple example of the larger problem I am trying to solve

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c( NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND

```{r}
is.na (x) & is.na(y)
```

Take the sum() to find out how many TRUE values we have and thus how many NAs we had in both x and y

```{r}
sum(is.na (x) & is.na(y))
```

Now I can make this into our first function...

```{r}
both_na <-  function(x, y) {
  sum(is.na (x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test, test, test

```{r}
 x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
3 == 3
3 == 2
3 != 2
```

```{r}
length(x)
length (y2)
```

Now lets try the both_na2() function on our different length input vectors

```{r}
#both_na2 (x, y2)
```

```{r}
which( c(F, F, T, F, T))

which( is.na( c(1, 2, NA, 4)))
```

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c( NA, 3, NA, 3, 4)

both_na3 (x, y)
```

## Intersect function

```{r}
df1
```

```{r}
df2
```

Make thing simple

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect (x, y)
```

```{r}
x %in% y
which ( x %in% y )
```

```{r}
x [x %in% y]
```

```{r}
cbind ( x [ x %in% y],
        y [ y %in% x] )
```

Use the RStudio shortcut 'CODE > EXTRACT FUNCTION' to turn our snippet into a working function

```{r}
gene_intersect <- function(x, y) {
  cbind ( x [ x %in% y],
          y [ y %in% x] )
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1, df2, by = "IDs")
```

## Practice writing our own function

Want to make a function called grade() tat will calculate the mean score and drop the lowest single score

```{r}
student1  <-  c(100, 100, 100, 100, 100, 100, 100, 90)
student2  <-  c(100, 90, 90, 90, 90, 90, 97, 80)


 
grade <- function(x) {
  (sum(x) - min(x))/ (length(x)-1)
}
```

```{r}
grade(student1)
grade(student2)
```








